---
format: gfm
---

```{r include = FALSE}
knitr::opts_chunk$set(comment = "#>")
suppressPackageStartupMessages(library(spatialreg))
```

# spdgp

<!-- badges: start -->

[![CRAN](https://www.r-pkg.org/badges/version/spdgp)](https://CRAN.R-project.org/package=spdgp)
[![CRAN Release](https://www.r-pkg.org/badges/last-release/spdgp)](https://CRAN.R-project.org/package=spdgp)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-20b2aa.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![R-CMD-check](https://github.com/JosiahParry/spdgp/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JosiahParry/spdgp/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

*Spatial Data Generation Processes*

`{spdgp}` is an R port of the [`pysal`](https://pysal.org/) module [`dgp`](https://pysal.org/spreg/api.html#dgp) within [`spreg`](https://pysal.org/spreg) library. 

`spdgp` is designed around the [`spdep`](https://r-spatial.github.io/spdep/) package's `listw` object for representing spatial weights matrices. 

## Overview

Use `spdgp` to generate data for the following models: 

- OLS: `sim_ols()`
- Spatial Error Model (SEM): `sim_sem()` 
- Spatial Lag Model (SAR): `sim_sar()`
- Spatially Lagged X Model (SLX): `sim_slx()`
- Spatial Lagged X Error Model (SLX Error): `sim_slx_error()`
- Spatial Autoregressive Model with Autoregressive Errors (SARAR / SAC / "Combo" model): `sim_sarar()`
- Spatial Durbin Model: `sim_durbin()`
- General Nested Model (GNM): `sim_gns()`
- Matrix Exponential Spatial Lag Model (MESS): `sim_mess()` 


## Installation 

- Install from [CRAN](https://CRAN.R-project.org/package=spdgp) with:

``` r
install.packages("spdgp")
```

- Install from source code on [GitHub](https://github.com/JosiahParry/spdgp) with:

```r
if (!requireNamespace("pak")) {
    install.packages("pak")
}

pak::pak("josiahparry/spgdp")
```


## Basic usage: 

We first need to create a spatial weights matrix to simulate based off of: 

```{r}
library(spdgp)

set.seed(42)

n <- 50
listw <- sim_grid_listw(10, 5)
```

Next we can simulate our error term, x from our betas.
```{r}
# simulate error 
u <- make_error(n, method = "normal")

# simulate x values based on uniform distribution
x <- make_x(n, method = "uniform")

# create x's according to an intercept and beta value
x_beta <- make_xb(x, c(1, 5))
```

Next, we'll simulate the y and specify the autoregrssive parameter $\rho = 0.5$.

```{r}
# simulate y from error and the x_beta
y <- sim_sar(u, x_beta, listw, rho = 0.5)
```

Fit an SAR model using simulated data.

```{r}
library(spatialreg)

sar_mod <- lagsarlm(y ~ x$x_1, listw = listw)

summary(sar_mod)
```

In the model we can see that the estimate of `rho` is quite close to the specified value of `0.5`.

